Plots

### 2025-08-02 src/plots.Rmd
Change plots for b/w print per pj suggestion

```{r}
library(tidyverse)
library(dplyr)
library(ggrepel)
library(patchwork)
library(glue)
library(kableExtra)
```

@@

# Plot performance of models using lines and points
plot_perf <- function(
        df_long,
        f = calc_pccc,
        group = "cod_name",
        color = "model_name",
        text_scale = 1,
        title = NULL,
        title_size = 10,
        x_title = "Partial Chance Corrected Concordance (PCCC) / R = PCCC Range",
        y_title = NULL,
        text_mid_size = 2.25 * text_scale,
        text_mid_vjust = 2.5,
        text_start_size = 2.75 * text_scale,
        text_end_size = 3.5 * text_scale,
        text_hjust = 0.2,
        add_min_points = F,
        add_max_points = F,
        mid_points_size = 4,
        min_points_size = 4,
        max_points_size = 4,
        x_limits = c(0, 1.15),
        x_breaks = seq(0, 1, by = 0.1),
        x_expand = expansion(add = 0.2),
        x_color_labels = c(
            "G3" = "GPT-3.5 (G3)",
            "G4" = "GPT-4 (G4)",
            "ISV" = "InSilicoVA (ISV)",
            "I5" = "InterVA-5 (I5)"
        ),
        x_title_size = 9 * text_scale,
        y_title_size = 9 * text_scale,
        x_text_size = 8 * text_scale,
        y_text_size = 9 * text_scale,
        sort_group = T,
        legend_position = "bottom",
        legend_title = element_blank(),
        legend_margin = margin(c(-8, -2, -2, -2)),
        ...
    ) {

@@

    # Plot metric for each model
    out <- ggplot(df_plot, aes(x = metric, y = .data[[group]])) +
        geom_segment( # min max lines
            data = df_ends,
            aes(
                x = min_metric,
                xend = max_metric,
                y = .data[[group]],
                yend = .data[[group]]
            ),
            linewidth = 0.75
        ) +
        geom_point( # points
            shape = 21,
            color = "black",
            aes(fill = .data[[color]]),
            size = mid_points_size
        ) +
        geom_text( # min model text
            data = df_ends,
            aes(
                x = min_metric,
                y = .data[[group]],
                label = paste0(model_min, ": ", round(min_metric, 2))
            ),
            hjust = 1 + text_hjust,
            vjust = 0.5,
            size = text_start_size
        ) +
        geom_text( # max model text
            data = df_ends,
            aes(
                x = max_metric,
                y = .data[[group]],
                label = paste0(
                    model_max, ": ", round(max_metric, 2),
                    " (R=", round(group_range, 2), ")"
                )
            ),
            hjust = 0 - text_hjust,
            vjust = 0.5,
            size = text_end_size
        ) +
        geom_text_repel( # mid model text
            data = df_mids,
            aes(label = model_label),
            color = "gray60",
            vjust = text_mid_vjust,
            max.overlaps = Inf,
            size = text_mid_size
        ) +
        scale_fill_discrete(
            labels = x_color_labels
        ) +
        scale_x_continuous(
            limits = x_limits,
            breaks = x_breaks,
            expand = x_expand
        ) +
        labs(
            title = title,
            x = x_title,
            y = y_title
        ) +
        theme_minimal() +
        theme(
            legend.position = legend_position,
            legend.title = legend_title,
            legend.margin = legend_margin,
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            plot.title = element_text(size = title_size),
            axis.title.x = element_text(size = x_title_size),
            axis.title.y = element_text(size = y_title_size),
            axis.text.x = element_text(size = x_text_size),
            axis.text.y = element_text(size = y_text_size)
        )

@@

# Prep data
df_plot <- df_long %>%
    group_by(model_name) %>%
    summarize(
        pccc = calc_pccc(cod, model_cghr10, N = ncod_all),
        csmf = calc_csmf_acc(cod, model_cghr10)
    ) %>%
    arrange(pccc) %>%
    mutate(
        model_name = model_name %>%
            case_when(
                . == "G3" ~ "GPT-3.5",
                . == "G4" ~ "GPT-4",
                . == "ISV" ~ "InSilicoVA",
                . == "I5" ~ "InterVA-5",
                .default = .
            ) %>%
            factor(., levels = .)
    ) %>%
    pivot_longer(
        c(pccc, csmf),
        names_to = "metric",
        values_to = "metric_value"
    ) %>%
    mutate(
        metric = str_to_upper(metric) %>%
            str_replace(., "CSMF", "CSMF Accuracy") %>%
            factor(., levels = rev(c("PCCC", "CSMF Accuracy"))),
        metric_label = paste0(metric, "=", round(metric_value, 2)) %>%
            str_replace(., " Accuracy", "")
    )

# Plot data
fig_pccc_csmf <- ggplot(
    df_plot,
    aes(
        x = metric_value,
        y = model_name,
        fill = metric
    )
) +
    geom_bar(
        stat = "identity",
        position = position_dodge(),
        color = "black",
        width = 0.7
    ) +
    geom_text(
        data = df_plot,
        aes(
            label = round(metric_value, 2)
        ),
        position = position_dodge(width = 0.7),
        color = "black",
        fill = "white",
        size = 4.5,
        label.r = unit(0, "pt"),
        vjust = 0.5,
        hjust = -0.25
    ) +
    scale_x_continuous(
        limits = c(0, 1),
        breaks = seq(0, 1, 0.1)
    ) +
    labs(
        title = NULL,
        x = NULL,
        y = NULL
    ) +
    guides(
        fill = guide_legend(reverse = TRUE)
    ) +
    theme_minimal() +
    theme(
        legend.position = c(.875, .45),
        legend.title = element_blank(),
        legend.text = element_text(size = 12),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.title = element_text(size = 10),
        axis.title.x = element_text(size = 9),
        axis.title.y = element_text(size = 9),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12)
    )

@@

# Plot perf for pccc by age groups
fig_pccc <- plot_perf(
    df_long %>%
        mutate(
            age_label = case_when(
                age_name == "Adult"
                    ~ paste0(age_name, "\n12+ Years\n(n=", n_adult, ")"),
                age_name == "Child"
                    ~ paste0(age_name, "\n28 Days to 11 Years\n(n=", n_child, ")"),
                age_name == "Neonate"
                    ~ paste0(age_name, "\n<28 Days\n(n=", n_neo, ")"),
                .default = age_name
            ) %>%
                factor(., levels = unique(.) %>% sort %>% rev)
        ),
    calc_pccc,
    group = "age_label",
    text_mid_vjust = 2,
    sort_group = F,
    x_limits = c(0.2, 0.9),
    legend_margin = margin(-8, -2, -2, -2),
    x_expand = waiver(),
    text_scale = 1.25
)

>>>>

```{r}
library(tidyverse)
library(dplyr)
library(ggrepel)
library(ggpattern)
library(patchwork)
library(glue)
library(kableExtra)
```

@@

# Plot performance of models using lines and points
plot_perf <- function(
        df_long,
        f = calc_pccc,
        group = "cod_name",
        color = "model_name",
        text_scale = 1,
        title = NULL,
        title_size = 10,
        x_title = "Partial Chance Corrected Concordance (PCCC)",
        y_title = NULL,
        text_mid_size = 2.25 * text_scale,
        text_mid_vjust = 2.5,
        text_start_size = 2.75 * text_scale,
        text_end_size = 3.5 * text_scale,
        text_hjust = 0.2,
        add_min_points = F,
        add_max_points = F,
        mid_points_size = 4,
        min_points_size = 4,
        max_points_size = 4,
        x_limits = c(0, 1.15),
        x_breaks = seq(0, 1, by = 0.1),
        x_expand = expansion(add = 0.2),
        x_color_labels = c(
            "G3" = "GPT-3.5 (G3)",
            "G4" = "GPT-4 (G4)",
            "ISV" = "InSilicoVA (ISV)",
            "I5" = "InterVA-5 (I5)"
        ),
        x_shapes = 21:25,
        x_title_size = 9 * text_scale,
        y_title_size = 9 * text_scale,
        x_text_size = 8 * text_scale,
        y_text_size = 9 * text_scale,
        sort_group = T,
        legend_position = "bottom",
        legend_title = element_blank(),
        legend_margin = margin(c(-8, -2, -2, -50)),
        legend_add = c("R = Range"),
        ...
    ) {

@@

    # Add blank legend items
    if (!is.null(legend_add)) {

        # Add to manual color labels
        for (i in legend_add) {
            x_color_labels[i] <- i
        }

        # Add to factor levels
        df_plot[[color]] <- factor(
            df_plot[[color]],
            levels = c(levels(df_plot[[color]]), legend_add)
        )
    }

    # Plot metric for each model
    out <- ggplot(df_plot, aes(x = metric, y = .data[[group]])) +
        geom_segment( # min max lines
            data = df_ends,
            aes(
                x = min_metric,
                xend = max_metric,
                y = .data[[group]],
                yend = .data[[group]]
            ),
            linewidth = 0.75
        ) +
        geom_point( # points
            color = "black",
            aes(
                shape = .data[[color]],
                fill = .data[[color]]
            ),
            size = mid_points_size
        ) +
        geom_text( # min model text
            data = df_ends,
            aes(
                x = min_metric,
                y = .data[[group]],
                label = paste0(model_min, ": ", round(min_metric, 2))
            ),
            hjust = 1 + text_hjust,
            vjust = 0.5,
            size = text_start_size
        ) +
        geom_text( # max model text
            data = df_ends,
            aes(
                x = max_metric,
                y = .data[[group]],
                label = paste0(
                    model_max, ": ", round(max_metric, 2),
                    " (R=", round(group_range, 2), ")"
                )
            ),
            hjust = 0 - text_hjust,
            vjust = 0.5,
            size = text_end_size
        ) +
        geom_text_repel( # mid model text
            data = df_mids,
            aes(label = model_label),
            color = "gray60",
            vjust = text_mid_vjust,
            max.overlaps = Inf,
            size = text_mid_size
        ) +
        scale_shape_manual(
            values = x_shapes,
            labels = x_color_labels,
            drop = ifelse(is.null(legend_add), T, F)
        ) +
        scale_fill_discrete(
            labels = x_color_labels,
            drop = ifelse(is.null(legend_add), T, F)
        ) +
        scale_x_continuous(
            limits = x_limits,
            breaks = x_breaks,
            expand = x_expand
        ) +
        labs(
            title = title,
            x = x_title,
            y = y_title
        ) +
        theme_minimal() +
        theme(
            legend.position = legend_position,
            legend.title = legend_title,
            legend.margin = legend_margin,
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            plot.title = element_text(size = title_size),
            axis.title.x = element_text(size = x_title_size),
            axis.title.y = element_text(size = y_title_size),
            axis.text.x = element_text(size = x_text_size),
            axis.text.y = element_text(size = y_text_size)
        )

@@

# Prep data
df_plot <- df_long %>%
    group_by(model_name) %>%
    summarize(
        pccc = calc_pccc(cod, model_cghr10, N = ncod_all),
        csmf = calc_csmf_acc(cod, model_cghr10)
    ) %>%
    arrange(pccc) %>%
    mutate(
        model_name = model_name %>%
            case_when(
                . == "G3" ~ "GPT-3.5",
                . == "G4" ~ "GPT-4",
                . == "ISV" ~ "InSilicoVA",
                . == "I5" ~ "InterVA-5",
                .default = .
            ) %>%
            factor(., levels = .)
    ) %>%
    pivot_longer(
        c(pccc, csmf),
        names_to = "metric",
        values_to = "metric_value"
    ) %>%
    mutate(
        metric = str_to_upper(metric) %>%
            str_replace(., "CSMF", "CSMF Accuracy") %>%
            factor(., levels = rev(c("PCCC", "CSMF Accuracy"))),
        metric_label = paste0(metric, "=", round(metric_value, 2)) %>%
            str_replace(., " Accuracy", ""),
        metric_pattern = case_when(
            metric == "PCCC" ~ "stripe",
            metric == "CSMF Accuracy" ~ "none"
        )
    )

# Plot data
fig_pccc_csmf <- ggplot(
    df_plot,
    aes(
        x = metric_value,
        y = model_name,
        fill = metric
    )
) +
    geom_bar_pattern(
        aes(pattern = metric_pattern),
        stat = "identity",
        position = position_dodge(),
        color = "black",
        width = 0.7,
        pattern_fill = "black",
        pattern_angle = 45,
        pattern_spacing = 0.05,
        pattern_colour = "black"
    ) +
    geom_text(
        data = df_plot,
        aes(
            label = round(metric_value, 2)
        ),
        position = position_dodge(width = 0.7),
        color = "black",
        fill = "white",
        size = 4.5,
        label.r = unit(0, "pt"),
        vjust = 0.5,
        hjust = -0.25
    ) +
    scale_x_continuous(
        limits = c(0, 1),
        breaks = seq(0, 1, 0.1)
    ) +
    scale_pattern_manual(values = c(
        "stripe" = "stripe",
        "none" = "none"
    )) +
    labs(
        title = NULL,
        x = NULL,
        y = NULL
    ) +
    guides(
        pattern = "none",
        fill = guide_legend(
            reverse = TRUE,
            override.aes = list(
                pattern = c("stripe", "none")
            )
        )
    ) +
    theme_minimal() +
    theme(
        legend.position = "bottom",
        legend.margin = margin(c(-8, -2, -2, -2)),
        legend.title = element_blank(),
        legend.text = element_text(size = 9),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.title = element_text(size = 10),
        axis.title.x = element_text(size = 9),
        axis.title.y = element_text(size = 9),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12)
    )

@@

# Plot perf for pccc by age groups
fig_pccc <- plot_perf(
    df_long %>%
        mutate(
            age_label = case_when(
                age_name == "Adult"
                    ~ paste0(age_name, "\n12+ Years\n(n=", n_adult, ")"),
                age_name == "Child"
                    ~ paste0(age_name, "\n28 Days to 11 Years\n(n=", n_child, ")"),
                age_name == "Neonate"
                    ~ paste0(age_name, "\n<28 Days\n(n=", n_neo, ")"),
                .default = age_name
            ) %>%
                factor(., levels = unique(.) %>% sort %>% rev)
        ),
    calc_pccc,
    group = "age_label",
    text_mid_vjust = 2,
    sort_group = F,
    x_limits = c(0.2, 0.9),
    x_expand = waiver(),
    text_scale = 1.25
)
